---
title: "Notebook 05"
output: html_document
---

## Getting Started

Before running this notebook, select "Session > Restart R and Clear Output" in
the menu above to start a new R session. This will clear any old data sets and
give us a blank slate to start with.

After starting a new session, run the following code chunk to load the
libraries and data that we will be working with today.

```{r, include=FALSE, message=FALSE}
library(tidyverse)
library(ggrepel)
library(smodels)
library(stringi)

theme_set(theme_minimal())
options(dplyr.summarise.inform = FALSE)
options(width = 77L)

make_scopus_graph_data <- function(input)
{
  cites_sub <- semi_join(cites, input, by = c("id_citing" = "id"))
  cites_sub <- semi_join(cites_sub, input,  by = c("id_cited" = "id"))
  set.seed(1)
  graph <- sm_graph_layout(cites_sub)
  nodes <- inner_join(inner_join(scotus, graph$node, by = "id"), gr_all, by = "id")
  nodes$cluster_all <- as.character(nodes$cluster_all)
  edges <- graph$edge
  return(list(nodes = nodes, edges = edges))
}
```

## Supreme Court Data

The Supreme Court, the highest court in the United States, is the most powerful
part of the judiciary branch. Many stakeholders are interested
in learning about patterns in the Supreme Court. They pose questions such as:

- Are certain kinds of law decided during certain periods of time?
- Are contentious cases more central than others? In other words, are the
important cases the ones where judges agree or disagree?
- Which area of the law does the case reflect (according to the WUSTL categories)?
- Which cases are central to case law?
- Which cases are more likely to cited with each other?
- Which cases connect different areas of the law?

Let's read in our data.

```{r, message=FALSE}
scotus <- read_csv(file.path("data", "scotus_dh.csv"))
cites <- read_csv(file.path("data", "scotus_cites_dh.csv"))
gr_all <- read_csv(file.path("data", "scotus_gr_all_dh.csv"))
```

Let's take a look at the data.

- term: Term of the decision.
- chief: The chief justice during the case.
- win_ideology: The ideology of the decision (liberal vs conservative)
- win_party: If favorable, the decision aligns with the win_ideology
- issue: The number of the area assigned by Washington University in St. Louis (more info below).
- issue_area: The name of the category assigned by Washington University in St. Louis (more info below)

```{r, message=FALSE}
scotus
```
In order to build networks, we need an edge list. The first (id_citing) is the 
court case cited by the other court case (id_cited).

```{r, message=FALSE}
cites
```

Before we do some network analysis, let's just count how often each case is cited 
(i.e. what we call the "degree" in network analysis). So, we can see 1965-122 is
the most cited case. What is the case?

```{r}
counts <- count(cites, id_cited, sort = TRUE)
counts <- left_join(counts, scotus, by = c("id_cited" = "id"))
counts
```
### Selecting Issues

In order to identify areas of that law that cases address,
Washington University in St. Louis assigned each United States Supreme Court
case to a legal issue (ex. desegregation).
The [code book](http://scdb.wustl.edu/documentation.php?s=1) lists the corresponding
issue assigned [each issue number in the data set](https://github.com/introdh2016/response3_network/blob/master/ussc-31.zip). 


```{r}
scotus_sub <- filter(scotus, issue %in% c(20130)) # Change to take a look at an area.
gr <- make_scopus_graph_data(scotus_sub)
nodes <- gr$nodes
edges <- gr$edges
```

Let's take a look at code 20130: Civil Rights.

```{r}
nodes
```

### Plotting Metadata

Let's do some EDA with ggplot to better understand our data.
We are just going to focus on 20130.

Let's try to get a sense if cases are contentious by looking at the minor votes.

```{r}
ggplot(nodes) +
  geom_bar(aes(x = minor_votes))
```

Let's now take a look the ideology of the decision. 

```{r}
ggplot(nodes) +
  geom_bar(aes(x = win_party, fill = win_ideology), position = "dodge")
```


Finally, let's explore when these cases took place and under which court.

```{r}
ggplot(nodes) +
  geom_point(aes(x = term, y = minor_votes, color = chief), size = 2) +
  scale_color_viridis_d()
```

## Network Analysis

Now that we've done some EDA, let's explore networks. We are still just looking at civil rights cases. Our edge list connects a case to the case that it cited. 

```{r}
ggplot(nodes, aes(x, y)) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), alpha = 0.1, data = edges) +
  geom_point(fill = "black", size = 3.5, show.legend = FALSE, pch=21) +
  theme_void()
```

Let's add a label to better understand the network.

```{r}
ggplot(nodes, aes(x, y)) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), alpha = 0.1, data = edges) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_text_repel(aes(label = id), size = 2) +
  theme_void()
```

Let's say we want to begin to figure out what each of these cases are. We can!

```{r}
filter(nodes, id == "2017-020")
```

### Metadata

We can now use the metadata about the case to garner further insights. Remember
how we could see during EDA that the Burger court was a time when a lot of cases
were decided. We can incorporate that knowledge into our network. 

```{r}
ggplot(nodes, aes(x, y)) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), alpha = 0.1, data = edges) +
  geom_point(aes(fill = chief), size = 3.5, pch=21) +    # you can change the "fill = " to other variables
  theme_void()
```

### Centrality

Often we are very interested in the center of a network. This helps us identify
the most important node/ vertices in a network. In other words, it helps us see
which nodes are highly connected, which reveals that they have a close
relationship. 

#### Eigenvalue Centrality

```{r}
ggplot(nodes, aes(x, y)) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), alpha = 0.1, data = edges) +
  geom_point(aes(fill = eigen), size = 3.5, show.legend = FALSE, pch=21) +
  scale_fill_distiller(palette = "Spectral") +
  theme_void()
```

Now, let's figure out which of the court cases have the highest centrality. 
So, which cases are all those red and organ nodes?

```{r}
arrange(nodes, desc(eigen))
```


Now, let's look at the centrality if we were to look at all of the cases in our
data set. What this tells us is that the case represented by the red node is the
most cited from this cluster within the entire data set. 

```{r}
ggplot(nodes, aes(x, y)) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), alpha = 0.1, data = edges) +
  geom_point(aes(fill = eigen_all), size = 3.5, show.legend = FALSE, pch=21) +
  scale_fill_distiller(palette = "Spectral") +
  theme_void()
```

So, let's figure out the exact case.

```{r}
arrange(nodes, desc(eigen_all))
```

#### Betweeness Centrality

While it's great to understand which cases are central, we also often want to know 
which case gets us from one area to another. In other words, which cases connects
us to another area/ community.

So, let's see which cases is a bridge within our network.

```{r}
ggplot(nodes, aes(x, y)) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), alpha = 0.1, data = edges) +
  geom_point(aes(fill = between), size = 3.5, show.legend = FALSE, pch=21) +
  scale_fill_distiller(palette = "Spectral") +
  theme_void()
```

Let's figure out the cases.

```{r}
arrange(nodes, desc(between))
```

Now, let's see which cases are a bridge with all the cases. 

```{r}
ggplot(nodes, aes(x, y)) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), alpha = 0.1, data = edges) +
  geom_point(aes(fill = between_all), size = 3.5, show.legend = FALSE, pch=21) +
  scale_fill_distiller(palette = "Spectral") +
  theme_void()
```

```{r}
arrange(nodes, desc(between_all))
```


### Clusters

Let's now take a look at the cases that are very clustered together. This indicates
cases that are citing each other.

```{r}
ggplot(nodes, aes(x, y)) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), alpha = 0.1, data = edges) +
  geom_point(aes(fill = cluster), size = 3.5, pch=21) +
  theme_void()
```

Let's look at the cases that comprise each cluter.

```{r}
split(sprintf("%s (%d)", nodes$case_name, nodes$term), nodes$cluster)
```

Now, let's look at clusters in relation to all of the data. 

```{r}
ggplot(nodes, aes(x, y)) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), alpha = 0.1, data = edges) +
  geom_point(aes(fill = cluster_all), size = 3.5, pch=21) +
  theme_void()
```

Let's look at the exact cases.

```{r}
split(sprintf("%s (%d)", nodes$case_name, nodes$term), nodes$cluster)
```

## Other methods for picking cases

Now, you can do EDA and networks with the rest of the cases. 

### Important Overall Cases

This data will look at the entire data set.

```{r}
num_cases <- 50
scotus_sub <- left_join(scotus, gr_all, by = "id")
scotus_sub <- arrange(scotus_sub, desc(eigen_all))
scotus_sub <- filter(scotus_sub, component_all == 1)
scotus_sub <- slice_head(scotus_sub, n = num_cases)
gr <- make_scopus_graph_data(scotus_sub)
nodes <- gr$nodes
edges <- gr$edges
```

### Important Within Issue Area

This data will look at an issue_area. This aggregates all of the cases under an area
such as Civil Rights. 

```{r}
num_cases <- 50
scotus_sub <- left_join(scotus, gr_all, by = "id")
scotus_sub <- filter(scotus_sub, issue_area == "Civil Rights")
scotus_sub <- arrange(scotus_sub, desc(eigen_all))
scotus_sub <- filter(scotus_sub, component_all == 1)
scotus_sub <- slice_head(scotus_sub, n = num_cases)
gr <- make_scopus_graph_data(scotus_sub)
nodes <- gr$nodes
edges <- gr$edges
```
