---
title: "Notebook 10"
output: html_document
---

## Getting Started

Before running this notebook, select "Session > Restart R and Clear Output" in
the menu above to start a new R session. This will clear any old data sets and
give us a blank slate to start with.

After starting a new session, run the following code chunk to load the
libraries and data that we will be working with today.

```{r, include=FALSE, message=FALSE}
library(tidyverse)
library(ggrepel)
library(smodels)

theme_set(theme_minimal())
options(dplyr.summarise.inform = FALSE)
```

## Movies Data


For our data, we are going to look at the top 20 grossing films
for the last 20 years (2000-2019). We have several valuables, they include:

- genre: the least common attached to the film from IMDB (tends to be the most unique)
- mpa: Motion Picture Association rating
- runtime: duration of the film in minutes
- gross: amount of money in U.S. dollars per million
- rating_count: the number of people who rated the film on IMDB
- rating: IMDB rating, out of 10
- metacritic: rating form Metacritic, out of 100
- poster_brigthness: how bright the color is in the movie poster


So now, let's read in our data.


```{r, message=FALSE}
movies <- read_csv(file.path("data", "movies_top_20_2000_2019.csv"))
head(movies)
```


### One-variable plots

Let's take an initial look at one variable. Along with helping
us begin to understand our data, this also allows us to do some checks
to make sure our data looks accurate.

Let's explore the MPA ratings.

```{r}
ggplot(data = movies, aes(x = mpa)) +
  geom_bar()
```

Let's explore the film's by genre.

```{r}
ggplot(data = movies, aes(x = genre)) +
  geom_bar() +
  coord_flip()
```

Let's explore the film by IMDB ratings. (They are out of 10.)

```{r}
ggplot(data = movies, aes(x = rating)) +
  geom_bar()
```


### Two-variable plots

While it's nice to explore by one variable, often we want to look at variables
in relation to another. This allows us to start asking more complex questions
about the data as well as to see more nuanced patterns.


For each one, think about what you'd expect to see before making the plot.
Once you've made the graph, ask yourself:
- Does the graph support that initial expectation?
- Make you ask a different question/ set of questions?
- See a different pattern?

Let's explore how many people gave a rating on IMDB in relation to
the average rating from the community.

```{r}
ggplot(data = movies, aes(x = rating_count, y = rating)) +
  geom_point()
```

Let's explore how much the film made in relation to the IMDB rating.  

```{r}
ggplot(data = movies, aes(x = gross, y = rating)) +
  geom_point()
```

Let's explore the IMDB community's rating in relation to genre.  


```{r}
ggplot(data = movies, aes(x = rating, y = genre)) +
  geom_count() +
  scale_size_area(max_size = 4)
```

Let's explore the rating in relation to the genre.  

```{r}
ggplot(data = movies, aes(x = mpa, y = genre)) +
  geom_count() +
  scale_size_area(max_size = 4)
```

## Label subsets


An important part of data viz, it thinking through all the pieces that we can use to better
communicate our findings/arguments.  Labels are one way.

Let's take a look at specific films.

```{r}
movie_subset <- filter(
  movies,
  title %in% c("Black Panther", "The Lord of the Rings: The Return of the King", "The Martian")
)

ggplot(data = movies, aes(x = gross, y = rating)) +
  geom_point(color = "grey95") +
  geom_point(data = movie_subset, color = "salmon") +
  geom_text_repel(aes(label = title), data = movie_subset, color = "salmon")
```

Let's take a look at a random set of movies (this case 15).

```{r}
movie_subset <- slice_sample(movies, n = 15)

ggplot(data = movies, aes(x = gross, y = rating)) +
  geom_point(color = "grey95") +
  geom_point(data = movie_subset, color = "salmon") +
  geom_text_repel(aes(label = title), data = movie_subset, color = "salmon")
```


Let's take a look at the top 10 grossing films.


```{r}
movie_subset <- arrange(movies, desc(gross))
movie_subset <- slice_head(movie_subset, n = 10)

ggplot(data = movies, aes(x = gross, y = rating)) +
  geom_point(color = "grey95") +
  geom_point(data = movie_subset, color = "salmon") +
  geom_text_repel(aes(label = title), data = movie_subset, color = "salmon")
```


Let's take a look at the crime films.


```{r}
movie_subset <- filter(movies, genre == "Crime")

ggplot(data = movies, aes(x = gross, y = rating)) +
  geom_point(color = "grey95") +
  geom_point(data = movie_subset, color = "salmon") +
  geom_text_repel(aes(label = title), data = movie_subset, color = "salmon")
```
