---
title: "Notebook 04"
output: html_document
---

## Getting Started

Before running this notebook, select "Session > Restart R and Clear Output" in
the menu above to start a new R session. This will clear any old data sets and
give us a blank slate to start with.

After starting a new session, run the following code chunk to load the
libraries and data that we will be working with today.


```{r}
#install.packages("topicmodels")
```


```{r, include=FALSE, message=FALSE}
library(tidyverse)
library(ggrepel)
library(smodels)
library(stringi)
library(cleanNLP)
library(topicmodels)

theme_set(theme_minimal())
options(dplyr.summarise.inform = FALSE)
options(width = 77L)
```

## State of the Union Data

For this analysis we will be looking at the text from the State of the Union Addresses.
The corpus includes nearly every Address; you can read the text of any address through
the *American Presidency Project*'s website:

   https://www.presidency.ucsb.edu/documents/presidential-documents-archive-guidebook/annual-messages-congress-the-state-the-union

To start, let's read in the main table of data. Note: Do not try to open this table 
directly as it may cause RStudio to crash.

```{r, message = FALSE}
doc <- read_csv(file.path("data", "sotu_document.csv.gz"))
```

Here are the first few rows of the dataset. Notice that all of the text is put in a
single, final column.

```{r}
head(doc)
```

It will be useful to have just the metadata columns of the data, to allow us to easily
look at them and join them into other datasets without slowing down RStudio. We do this
with the following code:

```{r}
meta <- select(doc, -text)  # create new dataset that does not have the "text" variable
```

In order to do an analysis with the text data, we need to split apart the words and 
add additional linguistic information. This would be time-consuming to do by hand,
but we can quickly estimate this data using a pre-trained NLP model. (For more on NLP, 
start here: https://en.wikipedia.org/wiki/Natural_language_processing). The code to do
this in R is quite short and given the next block of code. The code is commented out
because it can take 20-30 minutes to finish running. Instead, we can load a saved
version of the data.

```{r, message = FALSE}
# cnlp_init_udpipe()
# token <- cnlp_annotate(doc)$token
````

Instead of waiting for R to process the tokenized text, let's read in the data that
would result from running the above code:

```{r, message = FALSE}
token <- read_csv(file.path("data", "sotu_token.csv.gz"))
token
```

This dataset has one row per token (work or punctionation mark), along with data 
representing information about the token. Specifically:

  - doc_id: an identifier that maps into the table `meta`
  - sid: counts the sentences in the text
  - tid: gives an unique identifier for each token
  - token: text of the token itself
  - lemma: lemmatized version of the token (e.g., remove capitalization, makes all
    nouns singular, all verbs in infinitive form), removes morphological forms
  - upos: a "universal" part of speech (POS) code; mostly maps onto commonly taught
    POS categories, such as ADJ (adjective), NOUN, PROPN (proper noun), and ADV (adverb)
  - xpos: a more fine grained set of POS tags
  - pct_doc: a number from 0-100 indicating the percentage into each document where
    a word occurs
    
We will see several ways of studying the texts using the table of tokens in the following
sections, which are broken down by method. Note that some of the code becomes a bit
involved; try to focus on the ideas of each method rather than worrying too much about
the specific functions being used.

## Frequencies 

One of the most straight-forward techniques for analyzing textual documents is to see 
how frequently a specific lemma is used in the text. For this corpus, we can show the
changes of these ratios over time to see if any interesting patterns emerge. Below
we have the code to do this for detecting periods; feel free to look at other lemmas,
which can be punctuation or individual words.

```{r}
lemma_to_count <- "."   # change this to look at different lemmas

temp <- group_by(token, doc_id)
temp <- summarize(temp, lemma_pct = mean(lemma == lemma_to_count) * 100)
temp <- left_join(temp, meta, by = "doc_id")
ggplot(temp, aes(year, lemma_pct)) +
  geom_line()
```
For example, we can see that the sentences get shorter over time. Can you offer a hypothesis for why?


Similarly, we can see what percentage of terms come from a specific POS tag.

Before we explore, why might we be interested in POS?  

```{r}
pos_to_count <- "ADJ"   # try: NOUN, PROPN, ADJ, ADV, CONJ

temp <- group_by(token, doc_id)
temp <- summarize(temp, pos_pct = mean(upos == pos_to_count) * 100)
temp <- left_join(temp, meta, by = "doc_id")
ggplot(temp, aes(year, pos_pct)) +
  geom_line()
```



## TF-IDF

Sometimes we want the data to tell us what terms are important rather than assuming. 
One method for doing this is term-frequency inverse document frequency (TF-IDF).
The following code uses this method to summarize the "most important" terms in each document.

```{r}
temp <- filter(token, upos %in% c("NOUN", "ADJ", "ADV"))   # only look at nouns, adjectives, and adverbes
temp <- sm_text_tfidf(temp)
temp <- group_by(temp, doc_id)
temp <- arrange(temp, desc(tfidf))
temp <- slice_head(temp, n = 7)                            # show top 7 terms
temp <- summarize(temp, top_terms = paste(token, collapse = "; "))
temp <- ungroup(temp)
temp <- left_join(temp, meta, by = "doc_id")
temp <- select(temp, short_label, top_terms)
temp
```
For example, we didn't include proper nouns because they will often be very unique. 
They risk obscuring our ability to see larger themes.

Make sure your screen is wide enough to show the results next to each short label. Feel free
to change the number of top terms or adjust the parts of speech used in the analysis. If you
see any interesting patterns, feel free to go back to the previous section and see if there 
are overall temporal patterns as well.

## Proper Nouns

Another similar approach to TF-IDF that works well for this corpus is to look at all of the
proper nouns that are the most common in each document. The code below takes the most common
proper nouns (after removing some that appear in a very large number of every addresses).

```{r}
words_remove <- c("Senate", "House", "Representatives", "Congress", "United", "States",
                  "Madam", "Speaker", "Mr.", "Vice", "President", "America", "Americans",
                  "Members", "U.S.", "Members", "Gentlemen")
temp <- filter(token, upos == "PROPN")
temp <- filter(temp, !(token %in% words_remove))
temp <- group_by(temp, doc_id)
temp <- arrange(mutate(temp, cnt = n()), desc(cnt))
temp <- slice_head(group_by(temp, doc_id), n = 7)
temp <- summarize(temp, top_terms = paste(token, collapse = "; "))
temp <- ungroup(temp)
temp <- left_join(temp, meta, by = "doc_id")
temp <- select(temp, short_label, top_terms)
temp
```

Notice that there are some mistakes, but that this technique does give a good high-level
description of what historical events are happening at the time of the address.

## KWiC

Sometimes we are interested in the way that a word is used, not just how often it is seen
or how unique it is in the dataset. Once way to do this is through the technique called
Keywords in Context (KWiC). The following code allows you to input a term and see all of
the instances where the term is used in the data. The instances are aligned to make it 
easy to see the different context of each term.

```{r}
query <- "immigrant"  # put any word you would like to explore here

doc_t <- doc
doc_t$text <- stri_replace_all(doc_t$text, " ", fixed = "\n")
temp <- stri_locate_all(stri_trans_tolower(doc_t$text), fixed = query)
temp <- tibble(doc_id = doc$doc_id, locs = lapply(temp, function(v) v[,1]))
temp <- unnest(temp, cols = locs)
temp <- filter(temp, !is.na(locs))
temp <- left_join(temp, doc_t, by = "doc_id")
temp$context <- stri_sub(temp$text, temp$locs - 30, temp$locs + 30)
temp <- select(temp, short_label, context)
temp <- slice_tail(temp, n = 1000)   # for safety; do not show more than 1000 results
temp <- sprintf("%-18s|  %s", temp$short_label, temp$context)
cat(temp, sep = "\n")
```

Try some different terms and see if you can see changes in their usage over time.

## Dunn Log-Likelihood

The Dunn Log-Likelihood, or G-squared, statistic is a way of measuring terms that
most strongly differentiate between two sets of documents. The algorithm is balancing
frequency (i.e. how often is the word used) in relation to signal (i.e. how indicative). 
Here, we will start by picking two presidents and seeing what terms most strongly 
differ between their  speeches. For reference, here are the "short" names of all 
the presidents in our data (usually last name, except when ambiguous):

```{r}
unique(meta$short_name)
```

And here is the code that compares two presidents. To start, let's look at Obama
and Kennedy:

```{r}
president_1 <- "Obama"
president_2 <- "Kennedy"

tf <- cnlp_utils_tf(filter(token, upos %in% c("NOUN", "ADJ", "ADV")))
cnt1 <- apply(tf[meta$short_name == president_1,], 2, sum)
cnt2 <- apply(tf[meta$short_name == president_2,], 2, sum)
index <- which(pmin(cnt1, cnt2) > 0)
cnt1 <- cnt1[index]
cnt2 <- cnt2[index]

tot1 <- sum(cnt1)
tot2 <- sum(cnt2)
e1 <- tot1 * (cnt1 + cnt2) / (tot1 + tot2)
e2 <- tot2 * (cnt1 + cnt2) / (tot1 + tot2)
ll <- 2 * (cnt1 * log(cnt1 / e1) + cnt2 * log(cnt2 / e2)) * sign(cnt1 * log(cnt1 / e1))
pval <- 1 - pchisq(abs(ll), 1)
scores <- matrix(round(as.numeric(ll), 2), ncol = 1)
rownames(scores) <- names(ll)
colnames(scores) <- "G-squared"
scores <- scores[order(scores[,1], decreasing = TRUE),,drop=FALSE]
rbind(head(scores, 15), tail(scores, 15))
```

At the top are the 15 terms most frequently associated with president one and the bottom are
the 15 terms associated most strongly with president two. You can modify the selected presidents,
the parts of speech, and choose to show more terms.

Using a similar approach, we can identify the biggest differences between the start and end of
a particular speech. The code below shows the terms most differentiating between the first half
and the second half of Obama's 2009 speech.

```{r}
speech_label <- "Obama, 2009"
pct_cut_off <- 50

this_id <- filter(meta, short_label == speech_label)$doc_id
temp <- filter(token, doc_id == this_id)
temp <- mutate(temp, pid = if_else(pct_doc <= pct_cut_off, "first", "second"))
tf <- cnlp_utils_tf(filter(temp, upos %in% c("NOUN", "ADJ", "ADV")), doc_var = "pid")
cnt1 <- apply(tf[rownames(tf) == "first",,drop = FALSE], 2, sum)
cnt2 <- apply(tf[rownames(tf) == "second",,drop = FALSE], 2, sum)
index <- which(pmin(cnt1, cnt2) > 0)
cnt1 <- cnt1[index]
cnt2 <- cnt2[index]

tot1 <- sum(cnt1)
tot2 <- sum(cnt2)
e1 <- tot1 * (cnt1 + cnt2) / (tot1 + tot2)
e2 <- tot2 * (cnt1 + cnt2) / (tot1 + tot2)
ll <- 2 * (cnt1 * log(cnt1 / e1) + cnt2 * log(cnt2 / e2)) * sign(cnt1 * log(cnt1 / e1))
pval <- 1 - pchisq(abs(ll), 1)
scores <- matrix(round(as.numeric(ll), 2), ncol = 1)
rownames(scores) <- names(ll)
colnames(scores) <- "G-squared"
scores <- scores[order(scores[,1], decreasing = TRUE),,drop=FALSE]
rbind(head(scores, 15), tail(scores, 15))
```

The earlier terms are at the top of the output and the later terms are at the
bottom. As before, play around with changing the selected speech and see what
comes out of the analysis.


## Topic Models

Another approach to studying a corpus of texts is to run a Topic Model on the
data. The code below runs a model using nouns adjectives and adverbs with
16 groups. It will take a few minutes to finish running.

Change 'k = 'to the number of topics that you want to select.


```{r}

num_topics <- 16  # select number of topics here

tf_matrix <- cnlp_utils_tf(
  filter(token, upos %in% c("NOUN", "ADJ", "ADV")),
  min_df = 0.01, max_df = 0.5
)
lda_model <- LDA(x = tf_matrix, k = num_topics, control = list(seed = 2811, verbose = 1)) 
```

Once the model finishes, we can extract the output and produce a new tidy dataset 
containing all of the information about the topics.

```{r}
terms <- lda_model@terms
beta <- lda_model@beta
gamma <- lda_model@gamma
top_k <- get_terms(lda_model, k = 5)

topics <- tibble(
  doc_id = rep(meta$doc_id, ncol(gamma)),
  topic_num = as.numeric(col(gamma)),
  prob = as.numeric(gamma)
)
topics$topic_name <- apply(top_k, 2, paste, collapse = "; ")[topics$topic_num]
topics <- left_join(topics, meta, by = "doc_id")
```

There is a strong temporal component to the topic models, which we can see visually
in the following graphic (the topics have been ordered from the "oldest" to the "newest"):

```{r}
temp <- filter(topics, prob > 0.01)
temp <- mutate(group_by(temp, topic_name), my = mean(year))
temp <- arrange(ungroup(temp), my)
temp <- mutate(temp, topic_name = fct_inorder(topic_name))

ggplot(data = temp, aes(x = year, y = topic_name)) +
  geom_point(aes(size = prob, color = topic_name), show.legend = FALSE) +
  scale_size_area()
```

We can also look at the documents most closely associated with each topic:

```{r}
temp <- filter(topics, prob > 0.1)
temp <- mutate(group_by(temp, topic_name), my = mean(year))
temp <- arrange(ungroup(temp), my)
temp <- mutate(temp, topic_name = fct_inorder(topic_name))
temp$label <- sprintf("%s (%d%%)", temp$short_label, round(temp$prob * 100))
temp <- arrange(temp, desc(prob))
split(temp$label, temp$topic_name)
```

As well as a further analysis of the words that are most closely associated with each
topic:

```{r}
get_terms(lda_model, k = 20)  #change the number to see more or less words
```

You can use the topic modelling code in the first block to re-run the model with
a different number of topics and see how it changes the output.

## Principal Component Analysis

As a final method, we will use a technique called principal component analysis (PCA) to
show the relative "closeness" between the terms used in each document. The output
of the analysis can be used to visualize the relationship between each document.
To avoid creating too many labels, we will label only each president's first speech:

```{r}
tfidf <- cnlp_utils_tfidf(
  filter(token, upos %in% c("NOUN", "ADJ", "ADV")),
  min_df = 0.01, max_df = 0.5
)
temp <- mutate(sm_tidy_pca(tfidf), doc_id = as.numeric(document))
temp <- left_join(temp, meta, by = "doc_id")
temp_first <- temp[!duplicated(temp$short_name),]
ggplot(data = temp, aes(x = v1, y = v2)) +
  geom_point(aes(color = year), alpha = 0.3) +
  geom_text_repel(aes(label = short_label), data = temp_first, size = 3) +
  scale_color_viridis_c() +
  theme_void()
```

Note that the x- and y-axes have no specific meaning; what matters is how close
each dot is to one another.

You can modify the parts of speech codes used in the analysis, investigate the
overall temporal direction of the corpus, and see which presidents are most 
closely connected based on the words used in their texts. 

