---
title: "Notebook 01"
output: html_document
---

## R Notebooks

This file is called an R Notebook. It is a mixture of text (like this)
written in a format called *markdown*, and blocks of code that look
like this:

```{r}
print("hello world")
```

You can run code a block of code by click on the green arrow in the top-left
of the code block. Try this on the block above; you should see the result of
the addition show up as a result below the code.

Most of the notes this semester will be given in the form of a new notebook.
During class we will work through the notes and embedded questions. Depending
on the pace of the day, we may not finishing all of the questions in the
notebook. You should try to finishing the remaining questions for homework
before the next class meeting.

Today we will cover some of the basics of running R code and introducing the
various parts of the RCloud software.

## Running R Code

We now want to give a very brief overview of how to run R code. We
will now only show snippets of R code and the output rather than a
screen shot of the entire RStudio session. Though, know that you
should think of each of the snippets as occurring inside of one of
the grey boxes in an RMarkdown file.

In one of its most basic forms, R can be used as a fancy calculator. We already
saw this above. Or, for example, we can divide 12 by 4:

```{r}
12 / 4
```


## Assigning a Variable

We can also store values by creating new *objects* within R. To do
this, use the `<-` (arrow) symbol. For example, we can create a new
object called `mynum` with a value of `8` by:

```{r}
mynum <- 3 + 5
```

Notice that the number will also show up in the upper left hand corner
of the RStudio window. We can now use our new object `mynum` exactly the
same way that we we would use the number 8. For example, adding it to 1
to get the number nine:

```{r}
mynum + 1
```

## Running a Library

In general, we will be using libraries that will help us do our analysis.
This will give us access to data sets and functions. A function in R is 
a set of code that takes an input and provides an output. So

 ```{r}
library(stringi)
```

Stringi helps us work with strings. Often for us, this will be text, but 
remember that strings are just a set of characters. Let's saw we want to look
at word frequence. How might we do that?
 
 ```{r}
stri_split_boundaries("May the Force be with you.")
```
  

## Formatting

It is very important to properly format your code in a consistent way.
Even though the code may run without errors and produce the desired results,
you will make your life easier by writing well-formatted code from the start.
This makes it easier to read and debug in the future. We will follow the
following guidelines:

- always put one space before and after an equals sign or assignment arrow
- always put one space around mathematical operations (such as `+` and `*`)
- always put one space after a comma, but no space before a comma
- do not include any other additional spaces

In the code block below, I wrote some code that selects all of the
fruits and vegetables and produces a scatter plot with all the vegetables
in the data set. We will learn more about these functions code over the
next few weeks. For now, just reformat the code.

```{r}
mynum<-2+2+log(8,2)
mynum
```


Try running the code to see how easily RStudio makes it to embed
visualizations into RMarkdown notebooks.


## Practice: Running R Code

That's all we have today for new material. In the rest of the
notebook are some further practice questions to see how well you
understood the material.

In the code block below, make a variable named `fav_number` and set it
equal to your favorite number. (Note, the quotes are not part of the
variable name).

```{r}

```

In the code below, apply the function `sqrt` to your favorite
number. (Again, the quotes are not part of the function name).

```{r}

```

Note that the only point of the exercises here are to familiarize
you with running code in R, creating variables, and applying functions.
We will get to more interesting tasks soon!

