---
title: "Introduction to Data Visualization"
output: html_document
---

# Getting Started

Before running this notebook, select "Session > Restart R and Clear Output" in
the menu above to start a new R session. This will clear any old data sets and
give us a blank slate to start with.

After starting a new session, run the following code chunk to load the
libraries and data that we will be working with today.

```{r, include=FALSE, message=FALSE}
library(tidyverse)
library(ggrepel)

theme_set(theme_minimal())

movies <- read_csv(file.path("data", "movie_class.csv"))
```

I have set the options `include=FALSE` and `message=FALSE` to avoid cluttering
the solutions with all the output from this code.

# Data Vizualization

## EDA
In general, what we are going to be doing now is EDA, exploratory data analysis.
As you read in the Arnold and Tilton reading:

"While often not named, exploratory data analysis permeates the digital humanities.
EDA, described in 1977 in John Tukey’s seminal work by the same name, is a
subfield of statistics offering a conceptual structure for drawing information
from data sources. It often evokes new hypotheses and guides statisticians toward a
ppropriate techniques for further inferential modeling. Tukey championed these
techniques after becoming concerned that a substantial amount of statistical
analysis was conducted by blindly applying models without first considering
their appropriateness. His work, and its subsequent expansions over the past
forty-five years, helps lay out an approach for thinking critically about data
analysis in the humanities. EDA is concerned with exploring data by iterating
between summary statistics, modeling techniques, and visualizations."

 Assumptions/objectives:

 - an approach to analyzing data
 - about exploring the data to then formulate hypotheses
 - summarizing data to garner insights
 - determine if new/other data is needed

 So, how might we do this? We can use ggplot2!


## Grammar of Graphics

So, we are back to Arnold and Tilton:

"Tukey’s work on EDA called for constructing visualizations, such as
scatterplots, histograms, and box plots, to understand patterns within
multivariate datasets. The field of data visualization builds on his work,
treating visualizations as their own object of study.
Jacques Bertin’s Semiology of Graphics, William Cleveland’s texts
The Elements of Graphing Data and Visualizing Data, and Leland Wilkinson’s
The Grammar of Graphics are foundational works in statistical graphics.
Bertin argued that graphics “constitutes one of the basic sign-systems conceived
by the human mind for the purposes of storing, understanding, and communicating
essential information” (2). He showed explicitly how graphics can serve as their
own form of argument and evidence. Cleveland built on Bertin’s work by
theorizing a distinction between graphing, the process of visualizing raw data,
and fitting, the process of visualizing transformed data and statistical models.
Wilkinson extended Cleveland’s theory by distinguishing between the mathematical
abstraction of a graph and the physical manifestation of a rendered graphic.
He then set out to describe the fundamental units that comprised a visualization."

So now let's get a quick overview of ggplot2:
https://rfortherestofus.com/2019/07/a-short-overview-of-the-grammar-of-graphics/

For those who would like to learn more, watch Leland Wilkinson talk about
his work at an H20.ai event in NYC in 2019: https://www.youtube.com/watch?v=1X93Sum_SyM

### Elements

Data visualization is an incredibly important skill and a task that R is
particularly well-designed for. We are going to learn and use the **ggplot2**
package for building beautiful and informative graphics. The package makes it
easy to build fairly complex graphics in a way that is guided by a general
theory of data visualization. The only downside is that, because it is built
around a theoretical model rather than many one-off solutions for different
tasks, it has a steep initial learning curve. These notes will, hopefully,
make this as painless as possible.

The core idea of the grammar of graphics is that visualizations are composed
of independent *layers*. To describe a specific layer, we need to specify
several elements:

- *data*: the data set from which data will be taken to construct the plot
- *geom*: a description of what kinds of objects to plot (i.e., points, labels,
or boxes)  
- *aes*: a mapping from elements of the plot to columns in our data set (i.e.,
the position on the x-axis or the color of our points); it stands for aesthetics

You can describe virtually any type of visualization by putting together these
elements.

To show how to use the grammar of graphics, we will start by using the food
data set introduced in the previous notes, with each row describing a particular
item of food along with various nutritional information. The first plot we will
make is a scatter plot that investigates the relationship between calories and
the total fat (in grams) that are in a 100g portion of each food item. In the
language of the grammar of graphics we can describe this with the following
elements:

- *data*: our data set is called `movie_class`
- *geom*: we will build a plot with a points geometry; each row of data is
represented by a point
- *aes*: the x-axis will be associated with variables such as `budget` and the y-axis with
`run_time`

## Bar plots

Let's get started by making sure we know what our data is.

```{r}
movies
```

Now, let's start exploring. Let's start with a bar chart.
How many films won?

```{r}
ggplot(data = movies, aes(x = win)) +
  geom_bar()
```
We should know the answer! However, doing some basic checks helps us make sources
that our data is "clean" and that there aren't errors.

Let's look at country_of_origin. Look about right?

```{r}
ggplot(data = movies, aes(x = country_of_origin)) +
  geom_bar()
```

And how about genre? Any surprises?

```{r}
ggplot(data = movies, aes(x = genre)) +
  geom_bar()
```

Let's make this a bit easier to read!

```{r}
ggplot(data = movies, aes(x = fct_infreq(genre))) +   # order by frequency of genre
  geom_bar() +                                        # create bar plot
  coord_flip() +                                      # flip the axes
  labs(x = "Count", y = "Genre")                      # add axis labels
```

When we are doing EDA, we often won't add labels. We usually know pretty well
what the variables are. However, before circulating, we want to provide labels.

## Scatter plot

Often, we want not to just look at 1 variable, but want to look at the relationship
between two or more variables. This will open up more questions!

For example, might there be a relationship between a movie's budget and how much the
film earns at the box office?

```{r}
ggplot(data = movies, aes(x = boxoffice, y = budget)) +
  geom_point()
```

Any trends among the best picture winners?

```{r}
ggplot(data = movies, aes(x = budget, y = boxoffice, color = win)) +
  geom_point()
```

Can we see other patterns if we add film run time to our analysis?

```{r}
ggplot(data = movies, aes(x = budget, y = boxoffice, color = win, size = run_time)) +
  geom_point()
```

Can this data give us any insights into when to release the film if
a studio wants major box office returns?

```{r}
ggplot(movies, aes(x = date_released, y = boxoffice)) +
  geom_point()
```

## Labels

Before we share our findings, it's important to think about the information
that the viewer needs to interpret the visualization.

```{r}
ggplot(data = movies, aes(x = budget, y = boxoffice, label = movie_title)) +
  geom_text()
```

```{r}
ggplot(data = movies, aes(x = budget, y = boxoffice, label = movie_title)) +
  geom_text_repel()
```


## Highlighting subsets of the data

Often, we want to highlight a part of the data. We can do that by selecting a
subset and then plotting.

Let's take a look at the winners!

```{r}
movies_subset <- filter(movies, win %in% c("Y"))

ggplot(data = movies, aes(x = budget, y = boxoffice, label = movie_title)) +
  geom_point(color = "grey85") +
  geom_point(data = movies_subset, color = "olivedrab") +
  geom_text_repel(data = movies_subset, color = "olivedrab")
```

More colors! http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf


Let's take a look at the biographical films.


```{r}
movies_subset <- filter(movies, genre %in% c("biographical"))

ggplot(data = movies, aes(x = budget, y = boxoffice, label = movie_title)) +
  geom_point(color = "grey85") +
  geom_point(data = movies_subset, color = "lightpink1") +
  geom_text_repel(data = movies_subset, color = "lightpink1")
```


Let's take a random sample (say 15) and visualize them.

```{r}
movies_subset <- slice_sample(movies, n = 15)

ggplot(data = movies, aes(x = budget, y = boxoffice, label = movie_title)) +
  geom_point(color = "grey85") +
  geom_point(data = movies_subset, color = "goldenrod1") +
  geom_text_repel(data = movies_subset, color = "goldenrod1")
```
